---
title: "chl_processing"
author: "Kathleen Onorevole"
date: "March 2, 2017"
output: html_document
---

This is the processing script for the chlorophyll a data. This script reads in the raw data, adds habitat and age columns, sets factor levels, and exports a clean csv. There is an optional script for calculating the averages of each core type and exporting that csv (ie for PRIMER).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remember to set the working directory**

Load the packages needed for these analyses
```{r libraries, include = F}

library(plyr)
library(dplyr)

source("functions.R")
```

Read in the raw data
```{r read.in.data}

chl.raw <- read.csv("chlorophyll/raw/chla_all-seasons_all-sites.csv", header = T, stringsAsFactors = F)

```

Set factors for Season, Site, and Core Name. Add columns for age & set factors. Screen data for suspiciously high or low chlorophyll:phaeophytin ratio and eliminate faulty data.
Create 2 data frames based on habitats. First one creates habitat column that includes LoRef & HiRef. Second eliminates those samples and renames Ref as sandflat.
```{r order.variables}

chl.new <- set.factors(chl.raw)

#Excluding between oyster cores collected only in summer 2014 at IMS
BTO <- is.na(chl.new$CoreName)
chl.new <- chl.new[!BTO, ]

row.names(chl.new) <- NULL

#Screening data for samples with a high chlorophyll:phaeophytin ratio (should be ~ 1.0 - 1.7)
high.ratio <- which(chl.new$chl_phaeo_ratio > 1.7)
chl.new[high.ratio, ]
#Dropping Fall Carrot LO2, which has a chlorophyll:phaeophytin ratio of 4.1. This sample was already flagged for having a strange 665 absorbance, so something clearly went wrong and the data are not useful.
chl.drop <- chl.new[-68, ]
row.names(chl.drop) <- NULL

#Screening data for low chlorophyll:phaeophytin ratio.
low.ratio <- which(chl.drop$chl_phaeo_ratio < 1.2)
chl.drop[low.ratio, ]

#Dropping Fall IMS HO3, which has the lowest chl:phaeo ratio as well as (more concerningly) the only negative chlorophyll concentration. There was likely something off with this sample since all 3 of its parameters are outliers.
chl.drop <- chl.drop[-88, ]
row.names(chl.drop) <- NULL

chl.expanded <- add.age(chl.drop)
chl.expanded$Age <- factor(chl.expanded$Age, levels = c("0", "2", "7", "20"))

#Creating data frame that includes Ref, LoRef, & HiRef
chl.LoHiRef <- add.habitat(chl.expanded)
chl.LoHiRef$Habitat <- factor(chl.LoHiRef$Habitat,
                               levels = c("oyster", "marsh", "Ref", "LoRef", "HiRef"))

#Creating data frame without LoRef & HiRef
LoHiRef <- which(chl.expanded$CoreName == "LoRef" | chl.expanded$CoreName == "HiRef")
chl.noLoHi <- chl.expanded[-LoHiRef, ]
#Adding habitat column that uses "sandflat" instead of "Reference"
chl.sandflat <- add.sandflat(chl.noLoHi)
chl.sandflat$Habitat <- factor(chl.sandflat$Habitat, levels = c("oyster", "marsh", "sandflat"))

```

Write 2 csvs for clean data: one with all habitats, and one without LoRef or HiRef, and with Reference relabeled as Sandflat.
```{r clean.csv}

write.csv(chl.LoHiRef, "chlorophyll\\clean\\chl-clean-orig-habitats_all-seasons_all-sites.csv",
          row.names = F)

write.csv(chl.sandflat, "chlorophyll\\clean\\chl-clean-noLoHi_all-seasons_all-sites.csv", row.names = F)

```
