---
title: "Basic Graphs"
author: "Kathleen Onorevole"
date: "September 16, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document creates basic graphs of parameters, including changes across seasons. The graphs do NOT combine data sets- ie there are no graphs regressing one parameter against another. See the combo-graphs script for those types of figures.

Includes code for graphs of N2 flux, denitrification efficiency, N2O, oyster filtration, stem density. The end of the script includes code for graphs produced for the defense PPT and thesis.

**Remember to set working directory**

Load the packages needed for this analysis
```{r libraries}

library("plyr")
library("dplyr")
library("ggplot2")
library("ggpmisc")

source("functions.R")
```

``` {r load.data}

type.N2O <- read.csv("N2O\\clean\\N2O-by-type_all-seasons_all-sites.csv", header = T)

core.N2O <- read.csv("N2O\\clean\\N2O-by-core_all-seasons_all-sites.csv", header = T)

annual <- read.csv("combined-data\\clean\\annual-params_all-seasons_all-sites.csv", header = T)

DNF <- read.csv("N2-flux\\clean\\N2-flux-by-type_all-seasons_all-sites.csv", header = T)

SOD <- read.csv("SOD\\clean\\SOD-by-core_all-seasons_all-sites.csv", header = T)

#Set basic factors, which converts Lo/Hi Ref and Summer 2015 to NAs
DNF <- set.factors.basic(DNF)

#Eliminate Lo/Hi Ref and Summer 2015 data
DNF <- na.omit(DNF)

#Add columns for age and habitat and set factors
DNF <- add.age(DNF)
DNF$Age <- as.numeric(DNF$Age)
DNF <- add.habitat(DNF)
DNF$Habitat <- factor(DNF$Habitat, levels = c("oyster", "marsh", "Ref"))

filtration <- read.csv("oyster-stem\\raw\\oyster-filtration_all-seasons_all-sites.csv", header = T, stringsAsFactors = F)

stem.density <- read.csv("oyster-stem\\raw\\stem-density_annual_all-sites.csv", header = T, stringsAsFactors = F)

```

Preparing the N2 flux data for graphing. Includes eliminating Summer 2015 and Lo/Hi Ref, replacing reference labels with sandflat, setting factors correctly, and summarizing data for graphing purposes (ie standard error).
``` {r DNF.prep}

core.DNF <- read.csv("N2-flux\\clean\\N2-flux-by-core_with-summer15.csv", header = T)

#Eliminate Summer 2015 and Lo/Hi Ref data
core.DNF <- core.DNF[!core.DNF$Season == "Summer15", ]
core.DNF <- core.DNF[!core.DNF$CoreName == "LoRef", ]
core.DNF <- core.DNF[!core.DNF$CoreName == "HiRef", ]

#Reset the habitat levels to allow for sandflat
core.DNF$Habitat <- factor(core.DNF$Habitat, levels = c("oyster", "marsh", "Ref", "sandflat"))

#Create list of rows with Ref habitats and then replace that with sandflat
sub1 <- which(core.DNF$Habitat == "Ref")
core.DNF$Habitat[sub1] <- "sandflat"

#Create new data frame with stats relevant to graphing
avg_DNF <- ddply(core.DNF, .(Site, Season, Habitat), summarise,
                 avg_DNF = mean(N2Flux),
                 sd_DNF = sd(N2Flux),
                 N = length(N2Flux),
                 se_DNF = sd_DNF / N)

#Add age and adjust all the factor levels
avg_DNF <- add.age(avg_DNF)
avg_DNF$Age <- as.numeric(avg_DNF$Age)
avg_DNF$Season <- factor(avg_DNF$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
avg_DNF$Site <- factor(avg_DNF$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
avg_DNF$Habitat <- factor(avg_DNF$Habitat, levels = c("oyster", "marsh", "sandflat"))

```

Plotting bar graphs of denitrification across seasons and sites by habitat.
``` {r DNF.graphs}

avg_DNF %>%
  filter(Season == "Summer") %>%
  ggplot(aes(Habitat, avg_DNF, fill = Site)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = avg_DNF + se_DNF, ymin = avg_DNF - se_DNF), width = 0.3,
                position = position_dodge(0.9))

avg_DNF %>%
  filter(Site == "IMS") %>%
  ggplot(aes(Habitat, avg_DNF, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = avg_DNF + se_DNF, ymin = avg_DNF - se_DNF), width = 0.3,
                position = position_dodge(0.9)) +
  scale_fill_manual(values = c("darkgoldenrod1", "darkred", "darkblue", "darkgreen"))

avg_DNF %>%
  ggplot(aes(Habitat, avg_DNF, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = avg_DNF + se_DNF, ymin = avg_DNF - se_DNF), width = 0.3,
                position = position_dodge(0.9)) +
  ylab("Average N2 flux") +
  facet_grid(Site ~ .)

```

Preparing the denitrification efficiency data for graphing as above. Ultimately creating data frames with sets of average DNE by habitat by season by site, and an annual DNE by habitat by site.
``` {r efficiency.prep}

all.eff <- read.csv("DNE\\clean\\DNE-clean_no-summer15_all-sites.csv", header = T)

#Eliminate Summer 2015 and Lo/Hi Ref data
core.eff <- all.eff[!all.eff$CoreName == "LoRef", ]
core.eff <- core.eff[!core.eff$CoreName == "HiRef", ]

#Renumbers row names after subsetting so there aren't gaps
rownames(core.eff) <- NULL

#Assign 0 to Winter Carrot Reference 1 AS A PLACEHOLDER!!
#This does NOT mean 0 efficiency for real. This is the only set of cores that did not have a valid
#efficiency; ie there were negative DIN and N2 fluxes. In order for ggplot to leave a space in the
#bar graph, there needs to be some value there.
core.eff$DNE[159] <- 0

#Eliminate NAs, which were built into the clean data file
core.eff <- na.omit(core.eff)

core.eff <- set.factors.basic(core.eff)

#Add column for habitat with sandflat instead of Ref, and set factors
core.eff <- add.sandflat(core.eff)
core.eff$Habitat <- factor(core.eff$Habitat, levels = c("oyster", "marsh", "sandflat"))

#Calculate stats for graphing purposes
avg_eff <- ddply(core.eff, .(Site, Season, Habitat), summarise,
                 avg_eff = mean(DNE),
                 sd_eff = sd(DNE),
                 N = length(DNE),
                 se_eff = sd_eff / N)

annual_eff <- ddply(core.eff, .(Site, Habitat), summarise,
                    annual_eff = mean(DNE),
                    sd_eff = sd(DNE),
                    N = length(DNE),
                    se_eff = sd_eff / N)

```

Plotting bar graphs of DNF efficiency across seasons and sites by habitat. Also annual averages by habitat and site.
``` {r efficiency.graphs}

#Bar graph of DNE over seasons by habitat. Faceted by site.
avg_eff %>%
  ggplot(aes(Habitat, avg_eff, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = avg_eff + se_eff, ymin = avg_eff - se_eff), width = 0.3,
                position = position_dodge(0.9)) +
  ylab("Average denitrification efficiency (%)") +
  scale_fill_manual(values = c("darkgoldenrod1", "darkred", "darkblue", "darkgreen"), drop = F) +
  facet_grid(Site ~ .) 

#Bar graph of DNE over restored age by habitat.
annual_eff %>%
  ggplot(aes(Habitat, annual_eff, fill = Site)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = annual_eff + se_eff, ymin = annual_eff - se_eff), width = 0.3,
                position = position_dodge(0.9)) +
  ylab("Average denitrification efficiency (%)")

avg_DNF %>%
  filter(Site == "IMS") %>%
  ggplot(aes(Habitat, avg_DNF, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = avg_DNF + se_DNF, ymin = avg_DNF - se_DNF), width = 0.3,
                position = position_dodge(0.9)) +
  scale_fill_manual(values = c("darkgoldenrod1", "darkred", "darkblue", "darkgreen"))

```

Prepping N2O data for graphing through aggregation on various levels and setting factors appropriately.
``` {r N2O.prep}

#Calculate summary stats for graphing N2O
core.N2O <- add.sandflat(core.N2O)

N2O.avgs <- ddply(core.N2O, .(Season, Site, Habitat), summarise,
                  avg_N2O = mean(N2O_flux),
                  sd = sd(N2O_flux),
                  N = length(N2O_flux),
                  se = sd / N)

N2O.avgs$Season <- factor(N2O.avgs$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
N2O.avgs$Site <- factor(N2O.avgs$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
N2O.avgs$Habitat <- factor(N2O.avgs$Habitat, levels = c("oyster", "marsh", "sandflat"))

annual.N2O <- ddply(core.N2O, .(Site, Habitat), summarise,
                  annual_N2O = mean(N2O_flux),
                  sd = sd(N2O_flux),
                  N = length(N2O_flux),
                  se = sd / N)

annual.N2O$Site <- factor(annual.N2O$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
annual.N2O$Habitat <- factor(annual.N2O$Habitat, levels = c("oyster", "marsh", "sandflat"))

#Add column for habitat with sandflat instead of ref & set all factor levels
type.N2O <- add.sandflat(type.N2O)
type.N2O <- set.factors.basic(type.N2O)
type.N2O$Habitat <- factor(type.N2O$Habitat, levels = c("oyster", "marsh", "sandflat"))


```

Graphs visualizing N2O flux across seasons and annually.
``` {r N2O.graphs}

#Plot of N2O flux by season, with error bars
#Basically, flux is low
p1 <- ggplot(N2O.avgs, aes(Site, avg_N2O, fill = Habitat))
p1 + geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = avg_N2O - se, ymax = avg_N2O + se), position = position_dodge(0.9), width = 
                  0.3) +
  facet_grid(. ~ Season)

#Same plot, but without labeled (not lined up right) and without error bars
#Uses piping to group by habitat instead of calling on data frame with stats as p1
type.N2O %>%
  group_by(Habitat) %>%
  ggplot(aes(Site, avg_N2OFlux, fill = Habitat)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(aes(label = round(..y.., 2)), fun.y = mean, geom = "text", size = 3) +
  facet_grid(. ~ Season)

#Code for filtering by season and graphing N2O flux
N2O.avgs %>%
  group_by(Habitat) %>%
  filter(Season == "Summer") %>%
  ggplot(aes(Site, avg_N2O, fill = Habitat)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(aes(label = round(..y.., 2)), fun.y = mean, geom = "text", size = 3, vjust = -0.5)

#Annual averages of N2O flux, with error bars, color coded by habitat
p2 <- ggplot(annual.N2O, aes(Site, annual_N2O, fill = Habitat))
p2 + geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = annual_N2O - se, ymax = annual_N2O + se), position = position_dodge(0.9), width = 0.3)

```

Graphs of oyster filtration and stem grass density.
``` {r oyster.stem.graphs}

filtration$Site <- factor(filtration$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
filtration$Season <- factor(filtration$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
stem.density$Site <- factor(stem.density$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))

#Graph of oyster filtration by site, color coded by season
oyster.p1 <- ggplot(filtration, aes(Site, filtration, fill = Season))
oyster.p1 + geom_bar(stat = "identity", position = "dodge")

stem.p1 <- ggplot(stem.density, aes(Site, adj_density, fill = Site))
stem.p1 + stat_summary(fun.y = mean, geom = "bar", position = "dodge")

```

``` {r SOD.anovas}

hist(SOD$SOD)
Habitat.anova <- aov(SOD ~ Habitat, data = SOD)
summary(Habitat.anova)

Site.anova <- aov(SOD ~ Site, data = SOD)
summary(Site.anova)

all.SOD.anova <- aov(SOD ~ Season * Site * Habitat, data = SOD)
summary(all.SOD.anova)

```

Used to create figures for defense. Generally same or similar to figures created above, but with improved aesthetics for PPT presentation.
``` {r defense.figures}

avg_DNF_years <- mutate(avg_DNF, Years = ifelse(Site == "IMS" , "0 years",
                         ifelse(Site == "Carrot", "2 years",
                                 ifelse(Site == "NOAA", "7 years", "20 years"))))

avg_DNF_years$Years <- factor(avg_DNF_years$Years, levels = c("0 years", "2 years", "7 years", "20 years"))

tiff("Avg annual N2 fluxes", width = 300, height = 250, units = "mm", res = 300, compression = "lzw")

avg_DNF_years %>%
  ggplot(aes(Habitat, avg_DNF, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = avg_DNF + se_DNF, ymin = avg_DNF - se_DNF), width = 0.3,
                position = position_dodge(0.9)) +
  facet_grid(Years ~ .) +
  xlab(NULL) +
  ylab(expression(paste("Average ", N[2], " flux (", mu, "mol ", N[2], " ", m^-2, " ", h^-1, ")"))) +
  coord_cartesian(ylim = c(-10, 120)) +
  scale_y_continuous(breaks = seq(0, 125, 25)) +
  geom_hline(yintercept = 0, size = 0.5, colour = "black") +
  theme(panel.background = element_rect(fill = "grey97", colour = "black"),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 26, colour = "black"),
       axis.text.y = element_text(size = 12, colour = "black"),
       strip.text = element_text(size = 20),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  scale_fill_manual(values = c("darkgoldenrod1", "darkred", "darkblue", "darkgreen"),
                         labels = c("  Summer",
                                 "  Fall",
                                 "  Winter",
                                 "  Spring"))

dev.off()

annual_eff_years <- mutate(annual_eff, Years = ifelse(Site == "IMS" , "0 years",
                         ifelse(Site == "Carrot", "2 years",
                                 ifelse(Site == "NOAA", "7 years", "20 years"))))

annual_eff_years$Years <- factor(annual_eff_years$Years, levels = c("0 years", "2 years", "7 years", "20 years"))
annual_eff_years <- add.age(annual_eff_years)
annual_eff_years$Age <- as.numeric(annual_eff_years$Age)


#Bar graph of DNE over restored age by habitat.

tiff("Avg annual DNE", width = 300, height = 250, units = "mm", res = 300, compression = "lzw")

annual_eff_years %>%
  ggplot(aes(Habitat, annual_eff, fill = Years)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymax = annual_eff + se_eff, ymin = annual_eff - se_eff), width = 0.3,
                position = position_dodge(0.9)) +
  xlab(NULL) +
  ylab("Average denitrification efficiency (%)") +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major.y = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 26, colour = "black"),
       axis.text.y = element_text(size = 16, colour = "black"),
       strip.text = element_text(size = 20),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  guides(fill = guide_legend(title = "Restored Age")) +
  scale_fill_manual(values = c("slategray2", "skyblue2", "skyblue3", "skyblue4"),
                    labels = c("  0 years",
                               "  2 years",
                               "  7 years",
                               "  20 years")) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100))

dev.off()

#Bar graph of N2O flux over restored age by habitat.

annual.N2O.age <- add.age(annual.N2O)
annual.N2O.age$Age <- factor(annual.N2O.age$Age, levels = c("0", "2", "7", "20"))

tiff("Avg annual N2O", width = 300, height = 250, units = "mm", res = 300, compression = "lzw")

annual.N2O.age %>%
  ggplot(aes(Age, annual_N2O, fill = Habitat)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = annual_N2O - se, ymax = annual_N2O + se), position = position_dodge(0.9), width = 0.3) +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.title.x = element_text(margin = margin(t = 15), size = 26),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major.y = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 18, colour = "black"),
       axis.text.y = element_text(size = 18, colour = "black"),
       strip.text = element_text(size = 20),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
   xlab("Restored age (years)") +
   ylab(expression(paste("Average ", N[2], "O flux (", mu, "mol ", N[2], "O ", m^2, " ", h^-1, ")"))) +
   scale_fill_manual(values = c("darkorchid4", "olivedrab", "darkgoldenrod"),
                      labels = c("  oyster",
                                 "  marsh",
                                 "  sandflat")) +
  geom_hline(yintercept = 0, size = 0.8, colour = "black")

dev.off()

```

Used to create figures for thesis. Modeled after defense figures, but included additional data visualizations not relevant to defense.
**Successful execution of this code requires the data pre-processing above.
``` {r thesis.figures}

###Graph of N2O fluxes by habitat type for each site, faceted by season.

#Adding age as a factor to enable accurate x axis plotting.
N2O.avgs.age <- add.age(N2O.avgs)
N2O.avgs.age$Age <- factor(N2O.avgs.age$Age, levels = c("0", "2", "7", "20"))

#Plotting the graph per convention.

tiff("Figures\\Thesis\\Habitat-N2O-flux-by-Season", width = 300, height = 250, units = "mm", res = 300, compression = "lzw")

N2O.avgs.age %>%
  ggplot(aes(Age, avg_N2O, fill = Habitat))+
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = avg_N2O - se, ymax = avg_N2O + se), position = position_dodge(0.9), width = 
                  0.3) +
  facet_grid(. ~ Season) + 
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.title.x = element_text(margin = margin(t = 15), size = 26),
       axis.line.y = element_line(colour = "black", size = 0.8),
       axis.line.x = element_line(colour =  "black", size = 0.5),
       panel.grid.major.y = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 18, colour = "black"),
       axis.text.y = element_text(size = 18, colour = "black"),
       strip.text = element_text(size = 20),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
   geom_hline(yintercept = 0, size = 0.5, colour = "black") +
   xlab("Restored age (years)") +
   ylab(expression(paste("Average ", N[2], "O flux (", mu, "mol ", N[2], "O ", m^-2, " ", h^-1, ")"))) +
   scale_fill_manual(values = c("darkorchid4", "olivedrab", "darkgoldenrod"),
                      labels = c("  oyster",
                                 "  marsh",
                                 "  sandflat"))

dev.off()


###Graphing oyster filtration data

#Using function to add column with restored site age for x axis
oyster.age <- age.axis(filtration)

tiff("Figures\\Thesis\\Oyster-filtration-by-Season", width = 300, height = 250, units = "mm", res = 300, compression = "lzw")

oyster.age %>%
  ggplot(aes(Years, filtration, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") +
  xlab("Restored age") +
  ylab(expression(paste("Oyster filtration (L", "  ", m^-2, " ", h^-1, ")"))) +
  geom_hline(yintercept = 0, size = 0.5, colour = "black") +
  theme(panel.background = element_rect(fill = "grey97", colour = "black"),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.title.x = element_text(margin = margin(t = 15), size = 26),
       axis.line.y = element_line(colour = "black", size = 0.8),
       axis.line.x = element_line(colour = "gray 70", size = 0.5),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 12, colour = "black"),
       axis.text.y = element_text(size = 12, colour = "black"),
       strip.text = element_text(size = 20),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  geom_hline(yintercept = 0, size = 0.8, colour = "black") +
  scale_fill_manual(values = c("darkgoldenrod1", "darkred", "darkblue", "darkgreen"),
                         labels = c("  Summer",
                                 "  Fall",
                                 "  Winter",
                                 "  Spring"))

dev.off()


###Graphing S. alterniflora stem density

#Using function to add a column with restored site age for x axis
stem.density.age <- age.axis(stem.density)

tiff("Figures\\Thesis\\Stem-density-by-Site", width = 300, height = 250, units = "mm", res = 300, compression = "lzw")

stem.density.age %>%
  ggplot(aes(Years, adj_density)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge", fill = "darkgreen") +
  xlab("Restored age") +
  ylab(expression(paste(italic("S. alterniflora"), " density (stems", "  ", m^-2, ")"))) +
  theme(panel.background = element_rect(fill = "grey97", colour = "black"),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.title.x = element_text(margin = margin(t = 15), size = 26),
       axis.line.y = element_line(colour = "black", size = 0.8),
       axis.line.x = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 12, colour = "black"),
       axis.text.y = element_text(size = 12, colour = "black"),
       strip.text = element_text(size = 20))

dev.off()


```


``` {r idk}

annual %>%
  #filter(Site == "NOAA") %>%
  group_by(Season, CoreName) %>%
  ggplot(aes(CoreName, NH3flux, fill = Season)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(aes(label = round(..y.., 2)), fun.y = mean, geom = "text", size = 3) +
  facet_grid(Site ~ .)


```

OPTIONAL: writing csvs of data packaged above. Includes DNE
``` {r data.export}

write.csv(avg_eff, "combined-data\\clean\\DNE-prepped-for-graphing_all-seasons_all-sites.csv", row.names = F)

```

