---
title: "chlorophyll-analysis"
author: "Kathleen Onorevole"
date: "March 3, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document analyzes the chlorophyll a data collected as part of the oyster/marsh chronosequence.

**Remember to set working directory**

Load the packages needed for this analysis
```{r libraries}

library(plyr)
library(dplyr)
library(ggplot2)
library(ggpmisc)
library(car)

source("functions.R")
```

Read in data and order factors correctly. "All" includes Lo and Hi Ref cores; "basic" excludes them and refers to Reference as Sandflat in the Habitat column.
```{r load.data}

chl.all <- read.csv("chlorophyll//clean//chl-clean-orig-habitats_all-seasons_all-sites.csv", header = T)

chl.basic <- read.csv("chlorophyll//clean//chl-clean-noLoHi_all-seasons_all-sites.csv", header = T)

chl.all <- set.factors(chl.all)
chl.all$Habitat <- factor(chl.all$Habitat, levels = c("oyster", "marsh", "Ref", "LoRef", "HiRef"))

chl.basic$Season <- factor(chl.basic$Season, levels = c("Summer", "Fall", "Winter", "Spring", "Summer15"))
chl.basic$Site <- factor(chl.basic$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
chl.basic$CoreName <- factor(chl.basic$CoreName, levels = c("LO", "HO", "LM", "MM", "HM", "Ref"))

#Creating dataframes averaging the observations and calculating other summary stats useful for graphing

chl.all.avg <- ddply(chl.all, c("Season", "Site", "CoreName"), summarise,
                     avg.chl = mean(chla_conc),
                     N = length(chla_conc),
                     stdev = sd(chla_conc),
                     SE = stdev / sqrt(N))

chl.basic.avg <- ddply(chl.basic, c("Season", "Site", "CoreName"), summarise,
                     avg.chl = mean(chla_conc),
                     N = length(chla_conc),
                     stdev = sd(chla_conc),
                     SE = stdev / sqrt(N))

```

Checking whether chlorophyll data meet assumptions of ANOVA when grouped in various ways.
```{r assumptions}

#Eliminating Summer 2015 since that won't be included in the overall ANOVA
chl.main.avgs <- chl.basic.avg[!chl.basic.avg$Season == "Summer15", ]

#Creating model grouping by Season and checking normal distribution of RESIDUALS and heterosked
main.model <- lm(avg.chl ~ Season, data = chl.main.avgs)
main.res <- residuals(main.model)
qqnorm(residuals(main.model))
qqline(residuals(main.model))

#Passes normality OF RESIDUALS, passes heterosked grouped by CoreName, fails heterosked grouped by Season
#Using Levene Test because I've since read that it's preferred
shapiro.test(main.res)
leveneTest(avg.chl ~ Season, data = chl.main.avgs)
leveneTest(avg.chl ~ CoreName, data = chl.main.avgs)

plot(avg.chl ~ Season, data = chl.main.avgs)
plot(chl.main.avgs$Season, main.res, ylab = "Residuals", xlab = "Season")
plot(main.model$fitted.values, main.res, ylab = "Residuals", xlab = "Predicted")


#Create model grouping by Season and CoreName (proxy for elevation) -- This is a more realistic way to examine the data altho the "elevations" could be collapsed into habitats in future.
mainmod2 <- lm(avg.chl ~ Season + CoreName, data = chl.main.avgs)
main2res <- residuals(mainmod2)
qqnorm(main2res)
qqline(main2res)

#Checking for normality OF RESIDUALS (passes)
#Heterosked passes when looking at Season & Core Name as multiple independent variables... not sure what this means??
shapiro.test(main2res)
leveneTest(avg.chl ~ Season * CoreName, data = chl.main.avgs)

#Visualizing boxplots of chl grouped by CoreName & color-coded by Season (can swap)
chl.main.avgs %>%
  ggplot(aes(CoreName, avg.chl, colour = Season)) +
  geom_boxplot()

#ANOVA indicates that, when run as independent variables, only Season is significant
# ***This seems to be the most salient result of this chunk***
anova2 <- aov(avg.chl ~ Season + CoreName, data = chl.main.avgs)
summary(anova2)
TukeyHSD(anova2)


#Attempting the same process using Season & Habitat, mostly out of curiosity (this is probably not a statistically-sound way to go about looking for significance).

#First using my function to add a habitat column to the averaged core data.
chl.main.habitat <- add.habitat(chl.main.avgs)
chl.main.habitat$Habitat <- factor(chl.main.habitat$Habitat, levels = c("oyster", "marsh", "Ref"))

#Creating the linear model and visualizing a QQ plot
hab.model <- lm(avg.chl ~ Season * Habitat, data = chl.main.habitat)
hab.res <- residuals(hab.model)
qqnorm(hab.res)
qqline(hab.res)

#Passes both normality OF RESIDUALS and heterosked
shapiro.test(hab.res)
leveneTest(avg.chl ~ Season * Habitat, data = chl.main.habitat)

#ANOVA indicates sig diff only for Season
hab.anova <- aov(avg.chl ~ Season + Habitat, data = chl.main.habitat)
summary(hab.anova)
TukeyHSD(hab.anova)

site.model <- lm(avg.chl ~ Season * Site, data = chl.main.avgs)
resids.site <- residuals(site.model)
qqnorm(resids.site)
qqline(resids.site)

shapiro.test(resids.site)
leveneTest(avg.chl ~ Season * Site, data = chl.main.avgs)

site.anova <- aov(avg.chl ~ Season * Site, data = chl.main.avgs)
summary(site.anova)

#Exploring assumptions with all the replicates thrown in but excluding Summer 2015. This will probably not be the way to go, b/c chl is so variable that the replicates aren't true replicates.
reps.noS15 <- filter(chl.basic, !Season == "Summer15")

reps.model <- lm(chla_conc ~ Season + CoreName, data = reps.noS15)
qqnorm(residuals(reps.model))
qqline(residuals(reps.model))
shapiro.test(residuals(reps.model))

#### OLD CODE for assumption testing HIDDEN ####

#Checking all data first: Fails Bartlett test for Core Name, Site, Season; passes for Habitat
bartlett.test(chla_conc ~ CoreName, data = chl.all)
bartlett.test(chla_conc ~ Site, data = chl.all)
bartlett.test(chla_conc ~ Season, data = chl.all)
bartlett.test(chla_conc ~ Habitat, data = chl.all)
leveneTest(chla_conc ~ CoreName, data = chl.all)

bartlett.test(avg.chl ~ CoreName, data = chl.all.avg)
leveneTest(avg.chl ~ Site, data = chl.all.avg)

#Checking basic data: Fails Bartlett for all
bartlett.test(chla_conc ~ CoreName, data = chl.basic)
bartlett.test(chla_conc ~ Site, data = chl.basic)
bartlett.test(chla_conc ~ Season, data = chl.basic)
bartlett.test(chla_conc ~ Habitat, data = chl.basic)

#Modify below code to explore histograms for data sets, including and excluding Lo & Hi Ref
#Also interesting to explore phaeo conc and chl:phaeo ratio w histograms
#Habitat, Site, & Season look ok, CoreName is pretty spread out
chl.all %>%
  ggplot(aes(chla_conc)) +
  geom_histogram(stat = "bin", binwidth = 10) +
  facet_grid(CoreName ~ .)

chl.basic %>%
  ggplot(aes(phaeo_conc)) +
  geom_histogram(stat = "bin", binwidth = 10) +
  facet_grid(Site ~ .)

test.norm.all.CoreName <- chl.all %>%
  group_by(CoreName) %>%
  summarise(SW.pvalue = shapiro.test(chla_conc)$p.value) %>%
  mutate(Result = ifelse(SW.pvalue > 0.05, "Normal", "Non-Normal"))

test.norm.basic.CoreName <- chl.basic %>%
  group_by(CoreName) %>%
  summarise(SW.pvalue = shapiro.test(chla_conc)$p.value) %>%
  mutate(Result = ifelse(SW.pvalue > 0.5, "Normal", "Non-Normal"))

#Trying different transformations to meet assumptions
all.trans <- ddply(chl.all, c("Season", "Site", "CoreName", "CoreNo", "Habitat"), summarise,
                   sq_trans = chla_conc ^ 2,
                   sqrt_trans = chla_conc ^ (1/2),
                   ln_trans = log(chla_conc))

bartlett.test(sqrt_trans ~ Season, all.trans)
leveneTest(sqrt_trans ~ Season, all.trans)

chl.all.avg %>%
  ggplot(aes(avg.chl)) +
  geom_histogram(stat = "bin", binwidth = 10) +
  facet_grid(Site ~ .)

#Further exploring the ln transformation
log.fit <- lm(ln_trans ~ CoreName + Season, data = all.trans)

all.trans$resids <- residuals(log.fit)
all.trans$preds <- predict(log.fit)
plot(resids ~ preds, data = all.trans, xlab = "Predicted Values", ylab = "Residuals")

qqnorm(all.trans$resids)
qqline(all.trans$resids)

plot(all.trans$resids)
abline(0,0)

hist(all.trans$resids)

shapiro.test(all.trans$resids)
leveneTest(ln_trans ~ CoreName, data = all.trans)
leveneTest(ln_trans ~ Season, data = all.trans)

```

Create basic graphs of average chlorophyll.
```{r visualizing}

chl.main.avgs %>%
  ggplot(aes(Site, avg.chl, fill = Season)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge")

chl.main.habitat %>%
  ggplot(aes(Site, avg.chl, fill = Season)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  facet_grid(Habitat ~ .)

chl.all.avg %>%
  filter(Season == "Spring") %>%
  ggplot(aes(Site, avg.chl)) +
  geom_bar(aes(fill = CoreName), stat = "identity", position = "dodge")

#Using stat_summary to automatically calculate mean values to explore data
p1 <- ggplot(chl.all.avg, aes(CoreName, avg.chl, fill = Site))
p1 + stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.y = median, geom = "point", colour = "black", position = position_dodge(width = 0.9)) +
  facet_grid(Season ~ .)

p2 <- ggplot(chl.all.avg, aes(CoreName, avg.chl, fill = Site))
p2 + geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(ymin = avg.chl - SE, ymax = avg.chl + SE),
                width = 0.2, position = position_dodge(width = 0.9)) +
  facet_grid(Season ~ .)

chl.all.avg %>%
  filter(CoreName == "HO" | CoreName == "LO") %>%
  ggplot(aes(CoreName, avg.chl, fill = Season)) +
  geom_bar(stat = "identity", position = "dodge") #+
  #geom_errorbar(aes(ymin = avg.chl - SE, ymax = avg.chl + SE),
  #              width = 0.2, position = position_dodge(width = 0.9)) +
  #facet_grid(Season ~ .)

#LO vs HO pattern is interesting
chl.all %>%
  ggplot(aes(Habitat, chla_conc, fill = CoreName)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge")

#Value appear highest in fall; HO ~ Ref; no clear marsh elevation patterns
chl.all %>%
  ggplot(aes(Habitat, chla_conc, fill = CoreName)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge") +
  facet_grid(Season ~ .)

chl.all.avg %>%
  ggplot(aes(CoreName, avg.chl, fill = Site)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge")

chl.basic %>%
  ggplot(aes(Habitat, chla_conc, fill = CoreName)) +
  stat_summary(fun.y = mean, geom = "bar", position = "dodge")

chl.all.avg %>%
  filter(CoreName == "HM") %>%
  ggplot(aes(Site, avg.chl)) +
  geom_boxplot()

chl.all.avg %>%
  filter(Season == "Fall") %>%
  ggplot(aes(Site, avg.chl, fill = CoreName)) +
  geom_bar(stat = "identity", position = "dodge")


```

Comparing summers 2015 and 2016.
```{r summer.comparison}

#Creating new data frame with only Summer 2014 and 2015 data, and only LO, HO, and MM cores. Summer 2015 was abbreviated sampling with only 3 sets of cores collected at each location, so that's all that should be analyzed when considering the summer data.
chl.summer <- chl.all.avg %>%
  filter(Season == "Summer" | Season == "Summer15") %>%
  filter(CoreName == "LO" | CoreName == "HO" | CoreName == "MM")

#Summer 2015 has a compelling decrease from the youngest to oldest site, but it's probably not much ITO results, because summer 2014 does nearly the opposite...
ggplot(chl.summer, aes(Season, avg.chl, fill = Site)) +
  geom_bar(stat = "identity", position = "dodge")

#No systematic differences that would point to anything interesting going on in one season vs another. 
chl.summer %>%
  ggplot(aes(Season, avg.chl, fill = CoreName)) +
  facet_grid(Site ~ .) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_errorbar(aes(group = CoreName, ymin = avg.chl - SE, ymax = avg.chl + SE),
                width = 0.2, position = position_dodge(width = 0.9))

  

```
