---
title: "combo-graphs"
author: "Kathleen Onorevole"
date: "September 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document creates graphs of combined data sets. A running tally of what it includes: DNE in comparison to other site parameters, oyster filtration & marsh grass stem density, 

**Remember to set working directory**

Loads packages and functions needed for this analysis.
``` {r libraries}

library("plyr")
library("dplyr")
library("ggplot2")
library("ggpmisc")
library("ggrepel")

source("functions.R")
```

Read in csv data.
``` {r load.data}

annual <- read.csv("combined-data\\clean\\annual-params_all-seasons_all-sites.csv", header = T)

DNE <- read.csv("combined-data\\clean\\DNE-prepped-for-graphing_all-seasons_all-sites.csv", header = T)

filtration <- read.csv("oyster-stem\\raw\\oyster-filtration_all-seasons_all-sites.csv", header = T, stringsAsFactors = F)

stem.density <- read.csv("oyster-stem\\raw\\stem-density_annual_all-sites.csv", header = T, stringsAsFactors = F)

N2O <- read.csv("N2O\\clean\\N2O-by-type_all-seasons_all-sites.csv", header = T)

```

General data prep relevant to many combo graphs.  1) Change reference to sandflat and create dataframe with habitat averages; set factors. **This creates a new dataframe called annual.sandflat.** 2) Average each parameter by habitat.
``` {r general.prep}

#Convert Ref label to sandflat in annual dataset
annual.sandflat <- add.sandflat(annual)
annual.sandflat <- set.factors(annual.sandflat)
annual.sandflat$Habitat <- factor(annual.sandflat$Habitat, levels = c("oyster", "marsh", "sandflat"))

#Avering SOD, SOM, and inundation for each habitat type per season & site
habitat.avgs <- ddply(annual.sandflat, .(Season, Site, Habitat), summarise,
                    avg_N2flux = mean(N2flux),
                    avg_inund = mean(Inund),
                    avg_SOM = mean(SOM),
                    avg_SOD = mean(SOD),
                    avg_O2conc = mean(O2conc),
                    avg_NOxflux = mean(NOxflux),
                    avg_NH3 = mean(NH3flux))

habitat.avgs <- set.factors.habitat(habitat.avgs)

```

Prepare data for DNE analysis by combining the habitat average data frame with the one for DNE. Also creating a data frame with annual averages for DNE and other site parameters.
``` {r prep.DNE}

#Merging the averaged SOD with the pre-averaged DNE for graphing
DNE.merge <- merge(habitat.avgs, DNE, by = c("Season", "Site", "Habitat"), all.x = T, sort = F)

#Create data frame of annual averages
annual.DNE.merge <- ddply(DNE.merge, .(Site, Habitat), summarise,
                          annual_SOD = mean(avg_SOD),
                          annual_SOM = mean(avg_SOM),
                          annual_inund = mean(avg_inund),
                          annual_eff = mean(avg_eff),
                          annual_NH3 = mean(avg_NH3))

```

Graphing DNE in relationship to other factors, such as SOM, SOD, inundation, and NH3 flux. Graphs are on a seasonal and annual basis.
``` {r DNE.relationships}

#Formula for 2nd order polynomial equation
quad.formula <- y ~ poly(x,2)

#For SOD vs DNE by habitat, log relationship or 4th order polynomial is good, esp for sandflat
#Basically, DNE is not crashing when SOD increases, so it's not like you're burning out your system
#of DNF efficiency just b/c there's more microbial activity
p1 <- ggplot(DNE.merge, aes(avg_SOD, avg_eff, colour = Habitat))
p1 + geom_point() +
  geom_smooth(aes(group = Habitat), method = "lm", formula = y ~ log(x), se = F, size = 1) +
  stat_poly_eq(formula = y ~ log(x),
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ log(x)),
                  geom = "text",
                  aes(label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

#Relationship b/t SOD and DNE during summer only seems to suggest decrease for oysters,
#stable conditions for sandflat, parabola for marsh
DNE.merge %>%
  filter(Season == "Summer") %>%
  ggplot(aes(avg_SOD, avg_eff, colour = Habitat)) +
  geom_point()

#No solid relationship when SOM vs DNE divided by habitat
#For SOM vs DNE by Site, log relationship looks good: all but Carrot have p ~ = 0.06
p2 <- ggplot(DNE.merge, aes(avg_SOM, avg_eff, colour = Season))
p2 + geom_point() +
  geom_smooth(aes(group = Season), method = "lm", formula = y ~ log(x), se = F, size = 1) +
  stat_poly_eq(formula = y ~ log(x),
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ log(x)),
                  geom = "text",
                  aes(label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

#It does appear that there is a decrease in efficiency as SOM increases, at least in oysters,
#during the summer
DNE.merge %>%
  filter(Season == "Summer") %>%
  ggplot(aes(avg_SOM, avg_eff, colour = Habitat)) +
  geom_point()

#This is a more relevant question: as SOM increases, are you also increasing NH3 production?
#Because that would be an ecosystem disservice that we want to avoid, even if SOM is also increasing DNF
#This plot shows no- it's basically a flat line relationship as SOM increases
DNE.merge %>%
  ggplot(aes(avg_SOM, avg_NH3, colour = Habitat)) +
  geom_point()

#Same lack of result for SOD vs NH3
DNE.merge %>%
  ggplot(aes(avg_SOD, avg_NH3, colour = Site)) +
  geom_point()

#Also no real relationship between NH3 and SOM during the summer (which is more relevant b/c of the
#summer SOM/DNF relationship)
DNE.merge %>%
  filter(Season == "Summer") %>%
  ggplot(aes(avg_SOM, avg_NH3, colour = Habitat)) +
  geom_point()

#Increasing inundation seems to generally increase efficiency
p4 <- ggplot(DNE.merge, aes(avg_inund, avg_eff, colour = Season))
p4 + geom_point()

#On an annual basis, increasing SOD is NOT decreasing efficiency
p5 <- ggplot(annual.DNE.merge, aes(annual_SOD, annual_eff, colour = Habitat))
p5 + geom_point() +
  geom_smooth(aes(group = Habitat), method = "lm", formula = y ~ x, se = F, size = 1) +
  stat_poly_eq(formula = quad.formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

```

Preparing oyster filtration & marsh grass density data for graphing. Involves lots of averaging and creating data frames that match up correctly on varying levels of resolution. This is more limited than usual because the oyster and marsh physical info was only collected for one season.
``` {r prep.oyster.marsh}

filtration$Site <- factor(filtration$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
filtration$Season <- factor(filtration$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
stem.density$Site <- factor(stem.density$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
stem.density$CoreName <- factor(stem.density$CoreName, levels = c("LM", "MM", "HM"))

#### STARTING WITH OYSTER FILTRATION DATA ####

#Subsetting only data for oyster cores only from annual data set and resetting the row #s
annual.oyster <- annual[annual$Habitat == "oyster", ]
rownames(annual.oyster) <- NULL

#Creating seasonal averages of N2 flux for the oyster cores
#Added N column temporarily to check that it's calculating correctly-- probably need some kind of
#built-in checking system in future...
annual.oyster2 <- ddply(annual.oyster, .(Season, Site), summarise,
                       N2Flux = mean(N2flux),
                       avg_SOM = mean(SOM),
                       avg_SOD = mean(SOD),
                       NH3Flux = mean(NH3flux))

#Creating data frame with N2 flux and oyster filtration for each season at each site
filt.combo <- merge(x = filtration, y = annual.oyster2, by = c("Site", "Season"))

##### MARSH GRASS STEM DENSITY DATA ######

#Subset the marsh habitat cores and get rid of row #s
annual.marsh <- annual[annual$Habitat == "marsh", ]
rownames(annual.marsh) <- NULL

#Calculate annual average parameters at each site
annual.marsh2 <- ddply(annual.marsh, .(Site), summarise,
                       N2Flux = mean(N2flux),
                       marsh_SOM = mean(SOM),
                       marsh_O2 = mean(O2conc))

#Calculate average stem density at each site
stem.avgs <- ddply(stem.density, .(Site), summarise,
                   avg_density = mean(adj_density))

#Merge the stem density and other site parameter data frames by site
stem.combo <- merge(x = stem.avgs, y = annual.marsh2, by = "Site")

###### Below code calculates stem density and other site parameters including CORE NAME to explore the
###### different elevations involved

#Calculate the average IMS stem density (b/c there's only one elevation, MM, so it doesn't make sense
#to just have all the data duplicated)
IMS.avg <- stem.density %>%
  filter(Site == "IMS") %>%
  summarise(total = mean(adj_density))

#Assign the average to the first line of IMS data, and delete the other 2 lines
stem.density[1,3] <- IMS.avg
stem.density <- stem.density[-c(2,3), ]

#Calculate average stats based on core name at each site
annual.marsh.core <- ddply(annual.marsh, .(Site, CoreName), summarise,
                       N2Flux = mean(N2flux),
                       marsh_SOM = mean(SOM),
                       marsh_O2 = mean(O2conc))

#Combine stem density & other site parameters based on core name and site
stem.core.combo <- merge(x = stem.density, y = annual.marsh.core, by = c("Site", "CoreName"))

```

Graphs of oyster filtration and marsh grass density against other parameters.
``` {r oyster.marsh.relationships}

##### OYSTER FILTRATION ######

#Comparing relationship between oyster filtration and N2 flux
#Not much to go on really
filt.p1 <- ggplot(filt.combo, aes(filtration, N2Flux, colour = Site))
filt.p1 + geom_point(size = 6)

#Comparing NH3 flux to oyster filtration
#Highest NH3 emissions are associated with lowest filtrations- but there are very few points
filt.p2 <- ggplot(filt.combo, aes(filtration, NH3Flux, colour = Site))
filt.p2 + geom_point(size = 6)

#Filtration also doesn't seem linked to SOM
filt.p3 <- ggplot(filt.combo, aes(filtration, avg_SOM, colour = Site))
filt.p3 + geom_point(size = 8)

filt.p4 <- ggplot(filt.combo, aes(filtration, avg_SOD, colour = Site))
filt.p4 + geom_point(size = 8)

###### MARSH GRASS STEM DENSITY #####

#Density and DNF don't look like anything much- just b/c you increase the grass density,
#doesn't mean you're going to increase DNF (field of dreams restoration)
stem.p1 <- ggplot(stem.combo, aes(avg_density, N2Flux, colour = Site))
stem.p1 + geom_point(size = 6)

#But SOM and density are a straight line! Textbook! So other ecosystem services of restoration
stem.p2 <- ggplot(stem.combo, aes(avg_density, marsh_SOM, colour = Site))
stem.p2 + geom_point(size = 6) +
  geom_smooth(aes(group = 1), method = "lm", formula = y ~ x, se = F, size = 1) +
  stat_poly_eq(formula = y ~ x,
               aes(group = 1, label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(group = 1, label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

#Density has no clear relationship with [O2]
stem.p3 <- ggplot(stem.combo, aes(avg_density, marsh_O2, colour = Site))
stem.p3 + geom_point(size = 10)


#### Below graphs focus on the elevation as a potential driver of changes, since the marsh cores
#were collected at 3 different elevations (LM, MM, HM)

#No new trends when broken down by elevation, except that HM has a much higher density
stem.core.combo %>%
  ggplot(aes(adj_density, marsh_SOM, colour = Site)) +
  geom_point(size = 6) +
  facet_grid(CoreName ~ .)

#Including all elevations generally strengthens the relationship
stem.p4 <- ggplot(stem.core.combo, aes(adj_density, marsh_SOM, colour = CoreName))
stem.p4 + geom_point(size = 6) +
  geom_smooth(aes(group = 1), method = "lm", formula = y ~ x, se = F, size = 1) +
  stat_poly_eq(formula = y ~ x,
               aes(group = 1, label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(group = 1, label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

#Stem density by elevation vs N2 flux- no real trends, b/c Army is pretty much a flat line
stem.p5 <- ggplot(stem.core.combo, aes(adj_density, N2Flux, colour = Site))
stem.p5 + geom_point(size = 6)

```

Exploring relationship between N2 flux and other parameters: SOD, O2 concentration, nitroux oxide.
``` {r N2flux.relationships}

SOD.p1 <- ggplot(habitat.avgs, aes(avg_SOD, avg_N2flux, colour = Site))
SOD.p1 + geom_point(size = 4)

SOD.p2 <- ggplot(annual.sandflat, aes(SOD, N2flux, colour = Season))
SOD.p2 + geom_point(size = 4)

O2.p1 <- ggplot(annual.sandflat, aes(O2conc, N2flux, colour = Season))
O2.p1 + geom_point(size = 4)

```

``` {r prep.N2O}

N2O.merge <- merge(annual, N2O, by = c("Season", "Site", "CoreName"), all.x = T, sort = F)

N2O.merge <- add.sandflat(N2O.merge)
N2O.merge <- set.factors.habitat(N2O.merge)

```

``` {r N2O.relationships}

#Graphing N2 vs N2O by season shows fall/winter significant and distinct from summer/spring
N2O.p1 <- ggplot(N2O.merge, aes(N2flux, avg_N2OFlux, colour = Season))
N2O.p1 + geom_point(size = 4) +
  geom_smooth(aes(group = Season), method = "lm") +
  stat_poly_eq(formula = y ~ x,
               aes(group = Season, label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(group = Season, label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

#Graphing N2 vs N2O by site shows NOAA & Army significant but w/ opposite trends
N2O.p2 <- ggplot(N2O.merge, aes(N2flux, avg_N2OFlux, colour = Site))
N2O.p2 + geom_point(size = 4) +
  geom_smooth(aes(group = Site), method = "lm") +
  stat_poly_eq(formula = y ~ x,
               aes(group = Site, label = paste(..eq.label.., ..rr.label.., sep = "~~~")), parse = T) +
  stat_fit_glance(method = "lm",
                  method.args = list(formula = y ~ x),
                  geom = "text",
                  aes(group = Site, label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

#Using facet grid to illustrate N2 vs N2O across sites, color coded by season
N2O.p3 <- ggplot(N2O.merge, aes(N2flux, avg_N2OFlux, colour = Season))
N2O.p3 + geom_point(size = 4) +
  facet_grid(Site ~ .)

```

Creating polished figures for the thesis document. Note that some of the previous data wrangling (ie prep code chunks) is required to properly execute this code.
``` {r thesis.figs}

#Scatterplot of N2 flux against O2 concentration, color coded by season. This graph uses the annual.sandflat dataframe, which includes averages for each core type (ie LO), for each site & season.

tiff("Figures\\Thesis\\N2-flux-by-O2-conc", width = 300, height = 200, units = "mm", res = 300, compression = "lzw")

annual.sandflat %>%
  ggplot(aes(O2conc, N2flux, colour = Season)) +
  geom_hline(yintercept = 0, size = 0.2, colour = "grey23") +
  geom_vline(xintercept = 0, size = 0.2, colour = "grey23") +
  geom_point(size = 4) +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.x = element_text(margin = margin(t = 15), size = 26),
       axis.title.y = element_text(margin = margin(r = 15), size = 26),
       axis.line.x = element_line(colour = "grey23", size = 0.5),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 20, colour = "black"),
       axis.text.y = element_text(size = 20, colour = "black"),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  xlab(expression(paste(" Average ", O[2], " concentration (mg ", O[2], " ", L^-1, ")"))) +
  ylab(expression(paste("Average ", N[2], " flux (", mu, "mol N ", m^-2, " ", h^-1, ")"))) +
  geom_hline(yintercept = 0, size = 0.8, colour = "black") +
  #coord_cartesian(ylim = c(-10, 120)) +
  #scale_y_continuous(breaks = seq(-20, 120, 20)) +
  scale_colour_manual(values = c("darkgoldenrod1", "darkred", "darkblue", "darkgreen"),
                         labels = c("  Summer",
                                 "  Fall",
                                 "  Winter",
                                 "  Spring"))

dev.off()


#Scatterplot of DNE vs SOD, color coded by habitat. Each point is the average value for that parameter by season and habitat.

tiff("Figures\\Thesis\\DNE-vs-SOD-by-Habitat", width = 300, height = 200, units = "mm", res = 300, compression = "lzw")

DNE.merge %>%
  ggplot(aes(avg_SOD, avg_eff, colour = Habitat)) +
  geom_point(size = 4) +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.x = element_text(margin = margin(t = 15), size = 28),
       axis.title.y = element_text(margin = margin(r = 15), size = 28),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 20, colour = "black"),
       axis.text.y = element_text(size = 20, colour = "black"),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  scale_colour_manual(values = c("darkorchid4", "olivedrab", "darkgoldenrod"),
                      labels = c("  oyster",
                                 "  marsh",
                                 "  sandflat")) +
  xlab(expression(paste("Average SOD (", mu, "mol", " ", O[2], " ", m^-2, " ", h^-1, ")"))) +
  ylab("Average denitrification efficiency (%)") 

dev.off()


#Scatterplot of N2O vs N2 flux, color coded by Site with regression lines and equations. Also see code below the tiff chunk for supporting code.

#Add year descriptors instead of site names for the legend
N2O.merge2 <- age.axis(N2O.merge)

tiff("Figures\\Thesis\\N2O-vs-N2-by-Site", width = 300, height = 200, units = "mm", res = 300, compression = "lzw")

N2O.merge2 %>%
  ggplot(aes(N2flux, avg_N2OFlux, colour = Years)) +
  geom_point(size = 6) +
  geom_point(shape = 21, colour = "black", size = 6) +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.x = element_text(margin = margin(t = 15), size = 28),
       axis.title.y = element_text(margin = margin(r = 15), size = 28),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 20, colour = "black"),
       axis.text.y = element_text(size = 20, colour = "black"),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
   guides(colour = guide_legend(title = "Restored Age")) +
   scale_colour_manual(values = c("slategray2", "skyblue2", "skyblue3", "skyblue4"),
                      labels = c("  0 years",
                               "  2 years",
                               "  7 years",
                               "  20 years")) +
  xlab(expression(paste("Average ", N[2], " flux (", mu, "mol N ", m^-2, " ", h^-1, ")"))) +
  ylab(expression(paste("Average ", N[2], "O flux (", mu, "mol ", N[2], "O ", m^-2, " ", h^-1, ")"))) +
    coord_cartesian(xlim = c(0, 120), ylim = c(-0.6, 0.4)) +
  geom_smooth(aes(group = Years), method = "lm", se = F) +
  geom_text(data = reg.eqns, aes(x = x.coord, y = y.coord, label = label), hjust = 0, vjust = 0, colour = "black", parse = T)

dev.off()

#Supporting code for the above figure (N2O vs N2) is below

#Creates a separate data frame storing the relevant information for the geom_text label. This was recommended because it is supposed to plot better than typing everything into geom_text as individual commands. The x and y coords are the locations on the graph where the text will plot. The text lines are further lined up by including hjust and vjust in the geom_text command above. The labels are annotated with mathematical symbology, with text separated with ~. You need to include parse = T in geom_text command for this to read properly. See note at the end for differentiating by color.

reg.eqns <- data.frame(
  x.coord = c(80, 80, 80, 80),
  y.coord = c(-0.4, -0.45, -0.5, -0.55),
  label = c(paste("0~years:~R^2 == 0.021 ~~~ p == 0.591"),
            paste("2~years:~R^2 == 0.0084 ~~~ p == 0.670"),
            paste("7~years:~R^2 == 0.272 ~~~ p == 0.009"),
            paste("20~years:~R^2 == 0.341 ~~~ p == 0.0027"))
)

#To make the text colored too, use the following code:
#  type = c("color names", "color names") etc
# and colour = type in the geom_text command

#These should actually be run first, but they're kept down here for clarity. This creates the linear regression line that can be evaluated with summary() to get the R2 and p values.
IMS.fit <- lm(avg_N2OFlux ~ N2flux, data = subset(N2O.merge, Site == "IMS"))
Carrot.fit <- lm(avg_N2OFlux ~ N2flux, data = subset(N2O.merge, Site == "Carrot"))
NOAA.fit <- lm(avg_N2OFlux ~ N2flux, data = subset(N2O.merge, Site == "NOAA"))
Army.fit <- lm(avg_N2OFlux ~ N2flux, data = subset(N2O.merge, Site == "Army"))



#Scatterplot of N2 flux vs oyster filtration, color coded by site

#Add restored age for legend
filt.combo2 <- age.axis(filt.combo)

tiff("Figures\\Thesis\\N2-vs-filtration-by-Site", width = 300, height = 200, units = "mm", res = 300, compression = "lzw")

filt.combo2 %>%
  ggplot(aes(filtration, N2Flux, colour = Years)) +
  geom_point(size = 6) +
  geom_point(shape = 21, size = 6, colour = "black") +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.x = element_text(margin = margin(t = 15), size = 28),
       axis.title.y = element_text(margin = margin(r = 15), size = 28),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 20, colour = "black"),
       axis.text.y = element_text(size = 20, colour = "black"),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  guides(colour = guide_legend(title = "Restored Age")) +
  scale_colour_manual(values = c("slategray2", "skyblue2", "skyblue3", "skyblue4"),
                      labels = c("  0 years",
                               "  2 years",
                               "  7 years",
                               "  20 years")) +
  xlab(expression(paste("Oyster filtration (L ", m^-2, " ", h^-1, ")"))) +
  ylab(expression(paste("Average  ", N[2], " flux (", mu, "mol N ", m^-2, " ", h^-1, ")")))

dev.off()


#Scatterplot of N2 flux vs S alterniflora stem density, color coded by Site

#Adding restored age column for the legend
stem.core.combo2 <- age.axis(stem.core.combo)

tiff("Figures\\Thesis\\N2-vs-stem-density-by-Site", width = 300, height = 200, units = "mm", res = 300, compression = "lzw")

stem.core.combo2 %>%
  ggplot(aes(adj_density, N2Flux, colour = Years)) +
  geom_point(size = 6) +
  geom_point(shape = 21, size = 6, colour = "black") +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.x = element_text(margin = margin(t = 15), size = 28),
       axis.title.y = element_text(margin = margin(r = 15), size = 28),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 20, colour = "black"),
       axis.text.y = element_text(size = 20, colour = "black"),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  coord_cartesian(ylim = c(0,80)) +
  guides(colour = guide_legend(title = "Restored Age")) +
  scale_colour_manual(values = c("slategray2", "skyblue2", "skyblue3", "skyblue4"),
                      labels = c("  0 years",
                               "  2 years",
                               "  7 years",
                               "  20 years")) +
  xlab(expression(paste(italic("S. alterniflora"), " density (stems", " ", m^-2, ")"))) +
  ylab(expression(paste("Average  ", N[2], " flux (", mu, "mol N ", m^-2, " ", h^-1, ")")))

dev.off()


#Scatterplot of SOM vs stem density, color coded by age

#Add restored age column for legend
stem.combo2 <- age.axis(stem.combo)

tiff("Figures\\Thesis\\SOM-vs-stem-density-by-Site", width = 300, height = 200, units = "mm", res = 300, compression = "lzw")

stem.combo2 %>%
  ggplot(aes(avg_density, marsh_SOM, colour = Years)) +
  geom_point(size = 6) +
  geom_point(shape = 21, size = 6, colour = "black") +
  theme(panel.background = element_rect(fill = "white", colour = NA),
       axis.title.x = element_text(margin = margin(t = 15), size = 28),
       axis.title.y = element_text(margin = margin(r = 15), size = 28),
       axis.line.x = element_line(colour = "black", size = 0.8),
       axis.line.y = element_line(colour = "black", size = 0.8),
       panel.grid.major = element_line(colour = "grey70", size = 0.25),
       axis.text.x = element_text(size = 20, colour = "black"),
       axis.text.y = element_text(size = 20, colour = "black"),
       legend.title = element_text(size = 23, face = "bold"),
       legend.text = element_text(size = 20)) +
  coord_cartesian(ylim = c(0,5)) +
  guides(colour = guide_legend(title = "Restored Age")) +
  scale_colour_manual(values = c("slategray2", "skyblue2", "skyblue3", "skyblue4"),
                      labels = c("  0 years",
                               "  2 years",
                               "  7 years",
                               "  20 years")) +
  geom_smooth(aes(group = 1), method = "lm", formula = y ~ x, se = F, colour = "navy") +
  geom_text(aes(x = 25, y = 4.5), hjust = 0, vjust = 0, colour = "black", size = 5, 
            label = paste("R^2 == 0.9 ~~~ p == 0.502"), parse = T) +
  xlab(expression(paste(italic("S. alterniflora"), " density (stems", " ", m^-2, ")"))) +
  ylab(expression(paste("Average SOM (%)")))

dev.off()

#Creating linear fit to obtain R2 and p values. Separated from above tiff code for clarity.
stem.SOM.fit <- lm(marsh_SOM ~ avg_density, data = stem.combo)

```


