---
title: "Main DNF Analysis"
author: "Kathleen Onorevole"
date: "August 18, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remember to set working directory**

Load the packages required for these analyses
```{r libraries}

library("plyr")
library("dplyr")
library("ggplot2")
library("ggpmisc")

```

Read in relevant data sets. These should all be taken from the CLEAN folders, which are housed in my master R Markdown folder. The raw data have been processed in separate R Markdown documents; see those files for details. Each parameter has its own folder within the parent folder.

Also source the functions script.
```{r read.in}

all.N2 <- read.csv("N2-flux\\clean\\N2-flux-by-core_with-summer15.csv", header = T, stringsAsFactors = F)

DNE <- read.csv("DNE\\clean\\DNE-clean_no-summer15_all-sites.csv", header = T, stringsAsFactors = F)

source("functions.R")

```

Set the distinguishing labels as factors and implementing the preferred order. Uses set.factors function written by me.
``` {r order.variables}

all.N2 <- set.factors(all.N2)

DNE <- set.factors.basic(DNE)
DNE <- na.omit(DNE)

```

Checking for normal distributions based on various groupings
``` {r DNF.normality}

test.norm.Season <- all.N2 %>%
  group_by(Season) %>%
  summarise(SW.pvalue = shapiro.test(N2Flux)$p.value) %>%
  mutate(Result = ifelse(SW.pvalue > 0.05, "Normal", "Non-Normal"))

all.N2 %>%
  ggplot(aes(N2Flux)) +
  geom_histogram(stat = "bin", binwidth = 5) +
  facet_grid(Site ~ .)

test.norm.Site <- all.N2 %>%
  group_by(Site) %>%
  summarise(SW.pvalue = shapiro.test(N2Flux)$p.value) %>%
  mutate(Result = ifelse(SW.pvalue > 0.05, "Normal", "Non-Normal"))

log.Site <- all.N2 %>%
  mutate(log_N2Flux = log(N2Flux + 40, 10))

test.norm.log.Site <- log.Site %>%
  group_by(Site) %>%
  summarise(SW.pvalue = shapiro.test(log_N2Flux)$p.value) %>%
  mutate(Result = ifelse(SW.pvalue < 0.05, "Normal", "Non-Normal"))

#Using add.habitat function to add a column with the habitat (ie oyster, marsh)
N2.hab <- add.habitat(all.N2)

#Set the factor levels for the new habitat column
N2.hab$Habitat <- factor(N2.hab$Habitat, levels = c("oyster", "marsh", "Ref", "LoRef", "HiRef"))

#Checking for normal distribution when grouped by Habitat
test.norm.Habitat <- N2.hab %>%
  group_by(Habitat) %>%
  summarise(SW.pvalue = shapiro.test(N2Flux)$p.value) %>%
  mutate(Result = ifelse(SW.pvalue < 0.05, "Normal", "Non-Normal"))

```

Running ANOVAs on the basic N2 flux 
```{r DNF.anovas}

season.anova <- aov(N2Flux ~ Season, data = N2.hab)
summary(season.anova)

habitat.anova <- aov(N2Flux ~ Habitat, data = N2.hab)
summary(habitat.anova)

site.anova <- aov(N2Flux ~ Site, data = N2.hab)
summary(site.anova)

TukeyHSD(site.anova)

#Run ANOVA treating Season & Habitat independently
anova.N2 <- aov(N2Flux ~ Season + Habitat, data = N2.hab)
summary(anova.N2)

TukeyHSD(anova.N2)

#Run ANOVA incorporating interaction between Season & Habitat
interaction.anova.N2 <- aov(N2Flux ~ Season * Habitat, data = N2.hab)
summary(interaction.anova.N2)

nested.anova <- aov(N2Flux ~ (Season + Habitat) / CoreName, data = N2.hab)
summary(nested.anova)
```

Graphing DNE.
``` {r DNE.graphs}

DNE %>%
  ggplot(aes(Site, DNE, fill = CoreName)) +
  geom_bar(stat = "identity", position = "dodge") + facet_grid(Season ~ .)

```

Creating linear correlations for 
``` {r linear.models}

#Eliminate Summer 2015 data points
no.summer2 <- N2.hab[!N2.hab$Season == "Summer15", ]

#Calculate summary stats based on core type
no.summer2 <- full.summary(no.summer2)
no.summer2 <- na.omit(no.summer2)

#Add columns for age and habitat, and set factors appropriately
N2.age <- add.age(no.summer2)
N2.age$Age <- as.numeric(N2.age$Age)
#N2.age$Age <- factor(N2.age$Age, levels = c("0", "2", "7", "20"))

N2.age <- add.habitat(N2.age)
N2.age$Habitat <- factor(N2.age$Habitat, levels = c("oyster", "marsh", "Ref"))

#Graphing relationship b/t age and N2 flux
formula <- y ~ x

p1 <- ggplot(N2.age, aes(Age, avg_N2Flux, colour = Habitat))
p1 + geom_point() +
  geom_smooth(method = "lm", formula = formula, se = F) +
  stat_poly_eq(formula = formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               parse = T) +
  stat_fit_glance(method = "lm",
                  geom = "text",
                  aes(label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

N2.age %>%
  ggplot(aes(Age, avg_N2Flux, colour = Habitat)) +
  geom_point() +
  facet_grid(Season ~ .) +
  geom_smooth(method = "lm", formula = formula, se = F) +
  stat_poly_eq(formula = formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               parse = T) +
  stat_fit_glance(method = "lm",
                  geom = "text",
                  aes(label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")

N2.age %>%
  ggplot(aes(Age, avg_N2Flux, colour = CoreName)) +
  geom_point() +
  facet_grid(Season ~ .) +
  geom_smooth(method = "lm", formula = formula, se = F) +
  stat_poly_eq(formula = formula,
               aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
               parse = T) +
  stat_fit_glance(method = "lm",
                  geom = "text",
                  aes(label = paste("P-value =", signif(..p.value.., digits = 4), sep = "")),
                  label.x.npc = "right")
  
```

