---
title: "Inundation"
author: "Kathleen Onorevole"
date: "August 22, 2016"
output: html_document
---

This is the processing script for the inundation data. Data were created by converting HOBO water level data to % inundation using NOAA tide records.  This script reads in the raw inundation data, adds habitat and age columns, sets factor levels, and exports a clean csv. There is an optional script for calculating the annual averages of each core type and exporting that csv (ie for PRIMER).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remember to set the working directory**

Load the packages needed for these analyses
```{r libraries, include = F}

library("plyr")
library("dplyr")

source("functions.R")
```

Read in the raw data
``` {r read.in.data}

inund <- read.csv("E:\\R Markdown\\Inundation\\raw\\inundation-raw_all-sites_all-seasons.csv", header = T, stringsAsFactors = F)

```


Add columns for age and habitat using functions. Note that age for Summer 2015 will be one year too young; adjust this manually in main script if using second summer values. Since they are normally eliminated, did not adjust in function script.
Set the distinguishing labels as factors and implementing the preferred order
``` {r order.variables}

inund <- add.age(inund)
inund$Age <- as.numeric(inund$Age)
inund <- add.habitat(inund)

inund$Season <- factor(inund$Season, levels = c("Summer", "Fall", "Winter", "Spring", "Summer15"))
inund$CoreName <- factor(inund$CoreName, levels = c("LO", "HO", "LM", "MM", "HM", "Ref"))
inund$Site <- factor(inund$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
inund$Habitat <- factor(inund$Habitat, levels = c("oyster", "marsh", "Ref"))

```

Write csv for clean data.
``` {r core.inundation}

write.csv(inund, "E:\\R Markdown\\Inundation\\clean\\inundation-clean_all-sites_all-seasons.csv", row.names = F)

```

OPTIONAL: calculate the ANNUAL average inundation by core TYPE (ie IMS LO).  This produces a csv file with simply the mean inundation parameters for the entire year, generally to be used in some other type of analysis (ie input into PRIMER).
``` {r annual.inundation}

annual.inund <- ddply(inund, c("Site", "CoreName"), summarise,
             avg_PercInun = mean(Perc_Inun))

write.csv(annual.inund, "E:\\R Markdown\\Inundation\\clean\\inundation-by-type_annual-avg_all-sites.csv", row.names = F)

```
