---
title: "N2O_processing"
author: "Kathleen Onorevole"
date: "September 16, 2016"
output: html_document
---

This is the processing script for the N2O (NITROUS OXIDE) concentration data.  It reads in the raw N2O data, removes lost cores & aberrant readings, sets factor levels, and exports a clean csv. There are optional scripts for calculating the N2O values for each core type and annual N2O conc values and exporting the csv's (ie for PRIMER).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remember to set working directory**

Load the packages needed for this analysis
```{r libraries}

library("plyr")
library("dplyr")

source("functions.R")
```

Read in the raw data
```{r read.in}

raw.noxide <- read.csv("N2O\\raw\\N2O-raw_no-summer15_all-sites.csv", header = T, stringsAsFactors = F)

```

Eliminate NAs for missing cores & aberrant readings
``` {r cleaning.up}

all.noxide <- na.omit(raw.noxide)

```

Set the distinguishing labels as factors and implementing the preferred order.
``` {r order.variables}

all.noxide$Site <- factor(all.noxide$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
all.noxide$Season <- factor(all.noxide$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
all.noxide$CoreName <- factor(all.noxide$CoreName, levels = c("LO", "HO", "LM", "MM", "HM", "Ref"))

all.noxide <- add.age(all.noxide)
all.noxide$Age <- as.numeric(all.noxide$Age)
all.noxide <- add.habitat(all.noxide)

#Getting rid of IMS LoRef & HiRef
all.noxide <- na.omit(all.noxide)

```

Write csv of clean N2O conc & flux data for each core.
``` {r export.csv}

write.csv(all.noxide, "N2O\\clean\\N2O-by-core_all-seasons_all-sites.csv", row.names = F)

```

OPTIONAL: calculate the average N2O flux for each core TYPE (ie LO rather than LO1).  This produces a csv file with simply the mean N2O flux for each core type, generally to be used in some other type of analysis (ie input into PRIMER).
``` {r type.N2Ofux}

type.N2O <- ddply(all.noxide, c("Season", "Site", "CoreName"), summarise,
                 avg_N2OFlux = mean(N2O_flux))

write.csv(type.N2O, "N2O\\clean\\N2O-by-type_all-seasons_all-sites.csv", row.names = F)

```

OPTIONAL part II: calculate the ANNUAL average N2O flux by core TYPE.  This produces a csv file with simply the mean N2O flux parameters for the entire year, generally to be used in some other type of analysis (ie input into PRIMER).
``` {r annual.N2Oflux}

annual.N2O <- ddply(type.N2O, c("Site", "CoreName"), summarise,
             annual_N2OFlux = mean(avg_N2OFlux))

write.csv(annual.N2O, "N2O\\clean\\N2O-by-type_annual-avgs_all-sites.csv", row.names = F)

```


