---
title: "Denitrification Efficiency"
author: "Kathleen Onorevole"
date: "August 23, 2016"
output: html_document
---

This is the processing script to calculate denitrification efficiency, which is based on the N2 and DIN fluxes, calculated separately.  It reads in the cleaned up N2 and nutrient flux data, combines the two data sets, sets factor levels, calculates DNE, and exports a clean csv.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remember to set the working directory**

Load the packages needed for these analyses
```{r libraries, include = F}

library("plyr")
library("dplyr")

source("functions.R")

```

Read in the raw data
``` {r read.in.data}

N2.flux <- read.csv("E:\\R Markdown\\DNE\\raw\\N2-flux-for-DNE_no-summer15_all-sites.csv", header = T, stringsAsFactors = F)

nut.flux <- read.csv("E:\\R Markdown\\DNE\\raw\\nut-fluxes-for-DNE_no-summer15_all-sites.csv", header = T, stringsAsFactors = F)

```

Combine the two data sets to enable DNE calculations. Create a new column for only positive DIN and N2 fluxes, which represent fluxes out of the sediment. Only fluxes out are used to calculate DNE.
``` {r combine.data}

combined.flux <- left_join(N2.flux, nut.flux, by = c("Season", "Site", "CoreName", "CoreNo"))

combined.flux$positive.DIN <- combined.flux$flux_DIN
combined.flux$positive.DIN[combined.flux$positive.DIN < 0] <- 0

combined.flux$positive.N2 <- combined.flux$N2Flux
combined.flux$positive.N2[combined.flux$positive.N2 <0] <- 0

```

Adds a new column to calculate DNE. Note that there will be some NaN ("not a number") values because in some cases, we're attempting to divide by 0.

Cleans up the data frame to include only identifying info and DNE.
``` {r DNE.calculations}

DNE.calcs <- combined.flux %>%
  mutate(
    DNE = (positive.N2 / (positive.N2 + positive.DIN) * 100)
  )

keeps <- c("Season", "Site", "CoreName", "CoreNo", "DNE")
DNE.final <- DNE.calcs[keeps]

```


Set the distinguishing labels as factors and implementing the preferred order
``` {r order.variables}

DNE.final$Season <- factor(DNE.final$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
DNE.final$CoreName <- factor(DNE.final$CoreName, levels = c("LO", "HO", "LM", "MM", "HM", "Ref", "LoRef", "HiRef"))
DNE.final$Site <- factor(DNE.final$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))

```

Write csv of DNE values.
``` {r export.DNE}

write.csv(DNE.final, "E:\\R Markdown\\DNE\\clean\\DNE-clean_no-summer15_all-sites.csv", row.names = F)

```

OPTIONAL: calculate the average DNE for each core TYPE (ie LO rather than LO1), after getting rid of the NA values.  This produces a csv file with simply the mean DNE for each core type, generally to be used in some other type of analysis (ie input into PRIMER).
``` {r type.DNE}

DNE.noNA <- na.omit(DNE.final)

type.DNE <- ddply(DNE.noNA, c("Season", "Site", "CoreName"), summarise,
                 avg_DNE = mean(DNE))

write.csv(type.DNE, "E:\\R Markdown\\DNE\\clean\\DNE-by-type_no-summer15_all-sites.csv", row.names = F)

```
