---
title: "Nutrient Concentrations"
author: "Kathleen Onorevole"
date: "August 22, 2016"
output: html_document
---

This is the processing script for the nutrient concentration data. This script reads in the raw inundation data, adds habitat and age columns, sets factor levels, and exports a clean csv.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Remember to set the working directory**

Load the packages needed for these analyses
```{r libraries, include = F}

library("plyr")
library("dplyr")

source("functions.R")
```

Read in the raw data and eliminate missing core (Winter Carrot HO2).
``` {r read.in.data}

nut.concs.raw <- read.csv("E:\\R Markdown\\nutrient-concs\\raw\\nutrient-concs-raw_all-seasons_all-sites.csv", header = T, stringsAsFactors = F)

nut.concs.raw <- na.omit(nut.concs.raw)

```

Sets the readings that were below detection for each nutrient to 0 and recalculates ON.
``` {r non.detect}

nut.concs.raw$uM_NOx[nut.concs.raw$uM_NOx < 0.0507] <- 0
nut.concs.raw$uM_NH3[nut.concs.raw$uM_NH3 < 0.23857] <- 0
nut.concs.raw$uM_PO4[nut.concs.raw$uM_PO4 < 0.01967] <- 0
nut.concs.raw$uM_TN[nut.concs.raw$uM_TN < 0.75] <- 0

nut.concs.detect <- nut.concs.raw %>%
  mutate(
    uM_ON = (uM_TN - (uM_NOx + uM_NH3))
    )

```


Add columns for age and habitat using functions. Set the distinguishing labels as factors and implement the preferred order.
``` {r order.variables}

nut.concs <- add.age(nut.concs.detect)
nut.concs$Age <- as.numeric(nut.concs$Age)
nut.concs <- add.habitat(nut.concs)

nut.concs$Season <- factor(nut.concs$Season, levels = c("Summer", "Fall", "Winter", "Spring"))
nut.concs$CoreName <- factor(nut.concs$CoreName, levels = c("In1", "In2", "LO", "HO", "LM", "MM", "HM", "Ref", "LoRef", "HiRef"))
nut.concs$Site <- factor(nut.concs$Site, levels = c("IMS", "Carrot", "NOAA", "Army"))
nut.concs$Habitat <- factor(nut.concs$Habitat, levels = c("oyster", "marsh", "Ref", "LoRef", "HiRef"))

```

Write csv for clean data.  Also writes csv for redoing the flux & DNE calculations after "below detection" concentration results were adjusted.
``` {r export.clean}

write.csv(nut.concs, "E:\\R Markdown\\nutrient-concs\\clean\\nut-concs-clean_all-seasons_all-sites.csv", row.names = F)

write.csv(nut.concs.detect, "E:\\MS Research\\Nutrients\\nut-concs-detect-clean_all-seasons_all-sites.csv", row.names = F)

```
